---
title: "Control Flow, Version Control, Git & GitHub with R Programming"
author: "Cathrine Kanini"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: cosmo
    highlight: tango
    code_folding: show
  pdf_document:
    toc: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

# Part 1: Control Flow in R Programming

Control flow structures allow us to control the execution of code based
on conditions and repetitions.

------------------------------------------------------------------------

## 1. Decision Making Structures

Decision making in programming allows us to control the flow of
execution based on specific conditions. In R, various decision-making
structures help us execute statements conditionally. These include:

### 1.1 if Statement

The `if` statement evaluates a condition and executes code only if the
condition is TRUE.

**Syntax:**

```{r eval=FALSE}
if (condition) {
  # execute this code
}
```

**Example:**

```{r}
a <- 76
b <- 67

if (a > b) {
  c <- a - b
  print("condition a > b is TRUE")
  print(paste("Difference between a, b is:", c))
}

if (a < b) {
  c <- a - b
  print("condition a < b is TRUE")
  print(paste("Difference between a, b is:", c))
}
```

**Key Points:**

-   Only executes when condition is TRUE
-   No alternative action if FALSE
-   Can have multiple independent if statements

------------------------------------------------------------------------

### 1.2 if-else Statement

The `if-else` statement provides an alternative action when the
condition is FALSE.

**Syntax:**

```{r eval=FALSE}
if (condition) {
  # execute if TRUE
} else {
  # execute if FALSE
}
```

**Example:**

```{r}
a <- 67
b <- 76

if (a > b) {
  c <- a - b
  print("condition a > b is TRUE")
  print(paste("Difference between a, b is:", c))
} else {
  c <- a - b
  print("condition a > b is FALSE")
  print(paste("Difference between a, b is:", c))
}
```

**Practical Example: Grade Classification**

```{r}
score <- 85

if (score >= 90) {
  grade <- "A"
} else {
  grade <- "B or lower"
}

print(paste("Your grade is:", grade))
```

------------------------------------------------------------------------

### 1.3 if-else-if Ladder

Multiple conditions are evaluated sequentially until one is TRUE.

**Syntax:**

```{r eval=FALSE}
if (condition1) {
  # execute if condition1 is TRUE
} else if (condition2) {
  # execute if condition2 is TRUE
} else {
  # execute if all conditions are FALSE
}
```

**Example:**

```{r}
a <- 67
b <- 76
c <- 99

if (a > b && b > c) {
  print("condition a > b > c is TRUE")
} else if (a < b && b > c) {
  print("condition a < b > c is TRUE")
} else if (a < b && b < c) {
  print("condition a < b < c is TRUE")
}
```

**Practical Example: Complete Grade System**

```{r}
score <- 78

if (score >= 90) {
  grade <- "A"
  message <- "Excellent!"
} else if (score >= 80) {
  grade <- "B"
  message <- "Very Good!"
} else if (score >= 70) {
  grade <- "C"
  message <- "Good!"
} else if (score >= 60) {
  grade <- "D"
  message <- "Pass"
} else {
  grade <- "F"
  message <- "Fail"
}

cat(sprintf("Score: %d\nGrade: %s\n%s\n", score, grade, message))
```

------------------------------------------------------------------------

### 1.4 Nested if-else Statement

An if-else statement inside another if-else statement.

**Syntax:**

```{r eval=FALSE}
if (parent_condition) {
  if (child_condition1) {
    # code
  } else {
    # code
  }
} else {
  if (child_condition2) {
    # code
  } else {
    # code
  }
}
```

**Example:**

```{r}
a <- 10
b <- 11

if (a == 10) {
  if (b == 10) {
    print("a:10 b:10")
  } else {
    print("a:10 b:11")
  }
} else {
  if (a == 11) {
    print("a:11 b:10")
  } else {
    print("a:11 b:11")
  }
}
```

**Practical Example: Login System**

```{r}
username <- "admin"
password <- "pass123"

if (username == "admin") {
  if (password == "pass123") {
    print("✓ Login successful!")
  } else {
    print("✗ Incorrect password")
  }
} else {
  if (username == "") {
    print("✗ Username cannot be empty")
  } else {
    print("✗ User not found")
  }
}
```

------------------------------------------------------------------------

### 1.5 switch Statement

Compares an expression against multiple cases and executes matching
code.

**Syntax:**

```{r eval=FALSE}
switch(expression, 
       case1 = result1,
       case2 = result2,
       default_result)
```

**Example with Numbers:**

```{r}
day_num <- 3

day_name <- switch(day_num,
                   "1" = "Monday",
                   "2" = "Tuesday",
                   "3" = "Wednesday",
                   "4" = "Thursday",
                   "5" = "Friday",
                   "6" = "Saturday",
                   "7" = "Sunday",
                   "Invalid day")

print(paste("Day", day_num, "is", day_name))
```

**Example with Characters:**

```{r}
operation <- "add"

result <- switch(operation,
                 "add" = 5 + 3,
                 "subtract" = 5 - 3,
                 "multiply" = 5 * 3,
                 "divide" = 5 / 3,
                 "Unknown operation")

print(paste("Result:", result))
```

------------------------------------------------------------------------

## 2. Loops in R

Loops allow repetitive execution of code blocks.

### 2.1 for Loop

Used when the number of iterations is known beforehand.

**Syntax:**

```{r eval=FALSE}
for (value in sequence) {
  # code to execute
}
```

**Example 1: Simple Sequence**

```{r}
for (val in 1:5) {
  print(val)
}
```

**Example 2: Days of the Week**

```{r}
week <- c('Sunday', 'Monday', 'Tuesday', 'Wednesday', 
          'Thursday', 'Friday', 'Saturday')

for (day in week) {
  print(day)
}
```

**Example 3: Loop on a List**

```{r}
my_list <- list(1, 2, 3, 4, 5)

for (i in seq_along(my_list)) {
  current_element <- my_list[[i]]
  print(paste("Element", i, "is:", current_element))
}
```

**Example 4: Loop on a Matrix**

```{r}
my_matrix <- matrix(1:9, nrow = 3)
print("Matrix:")
print(my_matrix)

cat("\nIterating through matrix:\n")
for (i in seq_len(nrow(my_matrix))) {
  for (j in seq_len(ncol(my_matrix))) {
    current_element <- my_matrix[i, j]
    cat(sprintf("Position [%d,%d] = %d\n", i, j, current_element))
  }
}
```

**Example 5: Loop on a Data Frame**

```{r}
my_dataframe <- data.frame(
  Name = c("Joy", "Juliya", "Boby", "Marry"),
  Age = c(40, 25, 19, 55),
  Gender = c("M", "F", "M", "F")
)

print("Data Frame:")
print(my_dataframe)

cat("\nIterating through rows:\n")
for (i in seq_len(nrow(my_dataframe))) {
  current_row <- my_dataframe[i, ]
  cat(sprintf("Row %d: %s, Age %d, Gender %s\n", 
              i, current_row$Name, current_row$Age, current_row$Gender))
}
```

------------------------------------------------------------------------

### 2.2 while Loop

Runs as long as a condition is TRUE. Number of iterations unknown
beforehand.

**Syntax:**

```{r eval=FALSE}
while (condition) {
  # code to execute
}
```

**Example 1: Display Numbers 1 to 5**

```{r}
val <- 1
while (val <= 5) {
  print(val)
  val <- val + 1
}
```

**Example 2: Calculate Factorial**

```{r}
n <- 5
factorial <- 1
i <- 1

while (i <= n) {
  factorial <- factorial * i
  i <- i + 1
}

print(paste("Factorial of", n, "is", factorial))
```

**Example 3: Input Validation (Simulation)**

```{r}
attempts <- 0
max_attempts <- 3
success <- FALSE

while (attempts < max_attempts && !success) {
  attempts <- attempts + 1
  # Simulate user input (would normally use readline())
  input_value <- sample(c(TRUE, FALSE), 1)
  
  if (input_value) {
    success <- TRUE
    print(paste("✓ Success on attempt", attempts))
  } else {
    print(paste("✗ Failed attempt", attempts))
  }
}

if (!success) {
  print("Maximum attempts reached!")
}
```

------------------------------------------------------------------------

### 2.3 repeat Loop

Executes indefinitely until explicitly stopped with `break`.

**Syntax:**

```{r eval=FALSE}
repeat {
  # code to execute
  
  if (condition) {
    break
  }
}
```

**Example 1: Display Numbers 1 to 5**

```{r}
val <- 1

repeat {
  print(val)
  val <- val + 1
  
  if (val > 5) {
    break
  }
}
```

**Example 2: Repeat Statement**

```{r}
i <- 0

repeat {
  print("Data 4 Data!")
  i <- i + 1
  
  if (i == 5) {
    break
  }
}
```

**Example 3: Menu System**

```{r}
counter <- 0

repeat {
  counter <- counter + 1
  
  # Simulate menu selection
  choice <- sample(1:4, 1)
  
  cat(sprintf("\nIteration %d - Choice: %d\n", counter, choice))
  
  if (choice == 1) {
    print("Option 1 selected")
  } else if (choice == 2) {
    print("Option 2 selected")
  } else if (choice == 3) {
    print("Option 3 selected - Exiting")
    break
  } else {
    print("Invalid option")
  }
  
  # Safety break after 10 iterations
  if (counter >= 10) {
    print("Max iterations reached")
    break
  }
}
```

------------------------------------------------------------------------

## 3. Loop Control Statements

### 3.1 break Statement

Exits the loop immediately.

```{r}
for (i in 1:10) {
  if (i == 6) {
    print("Breaking at 6")
    break
  }
  print(i)
}
```

### 3.2 next Statement

Skips the current iteration and continues with the next.

```{r}
for (i in 1:10) {
  if (i %% 2 == 0) {
    next  # Skip even numbers
  }
  print(paste("Odd number:", i))
}
```

------------------------------------------------------------------------

# Part 2: Version Control Systems

## What is Version Control?

A **Version Control System (VCS)** is a tool that tracks and manages
changes to source code over time. It enables:

-   **Change Tracking**: Record every modification to files
-   **Collaboration**: Multiple developers working simultaneously
-   **History Management**: Access to complete project evolution
-   **Reversion**: Ability to restore previous versions
-   **Branching**: Parallel development of features

------------------------------------------------------------------------

## Components of Version Control Systems

### Key Concepts

1.  **Repository**: Central storage for all project files and their
    complete history
2.  **Revision**: A specific saved version of files (identified by
    unique ID/hash)
3.  **Branch**: Separate line of development for features or fixes
4.  **Merging**: Combining changes from different branches
5.  **Commit**: Snapshot of changes at a specific point in time

------------------------------------------------------------------------

## Types of Version Control Systems

### 1. Local Version Control Systems

-   Operates entirely on local machine
-   All history stored in local database
-   **Pros**: Simple, no network needed
-   **Cons**: No collaboration, single point of failure

### 2. Centralized Version Control Systems (CVCS)

Examples: **SVN (Subversion)**, CVS

**Architecture:** - Single central server stores all files - Developers
check out files from central location - Changes committed directly to
central server

**Workflow:** 1. **Checkout/Update**: Get latest files from server 2.
**Make Changes**: Edit files locally 3. **Commit**: Save changes to
central server

**Pros:** - Easy to understand and manage - Fine-grained access
control - Everyone sees what others are doing

**Cons:** - Single point of failure (server down = no work) - Requires
network connection - Limited offline capabilities

### 3. Distributed Version Control Systems (DVCS)

Examples: **Git**, Mercurial, Bazaar

**Architecture:** - Every developer has full repository copy - Complete
history available locally - Changes shared through push/pull

**Workflow:** 1. **Commit**: Save changes to local repository 2.
**Push**: Upload changes to remote repository 3. **Pull/Fetch**:
Download changes from remote repository

**Pros:** - Fast operations (local) - Work offline - Multiple backup
copies - Flexible workflows - No single point of failure

**Cons:** - Steeper learning curve - More complex commands

------------------------------------------------------------------------

## Popular Version Control Systems

### 1. Git (Most Popular)

-   Created by Linus Torvalds (2005)
-   Distributed VCS
-   Fast, efficient, lightweight
-   Excellent branching and merging
-   Powers GitHub, GitLab, Bitbucket

### 2. Subversion (SVN)

-   Centralized VCS
-   Still widely used in enterprises
-   Simpler than Git for basic use
-   Good for binary files

### 3. Mercurial

-   Distributed VCS like Git
-   Simpler interface than Git
-   Good for large projects
-   Used by Facebook, Mozilla (historically)

### 4. CVS (Concurrent Versions System)

-   One of the earliest VCS
-   Centralized system
-   Legacy system (mostly replaced)
-   Historical importance

### 5. Bazaar

-   Developed by Canonical (Ubuntu creators)
-   Supports both centralized and distributed workflows
-   Beginner-friendly
-   No longer actively developed

------------------------------------------------------------------------

# Part 3: Git - The Complete Guide

## 1. Introduction to Git

**Git** is an open-source **distributed version control system**
designed for:

-   Tracking source code changes
-   Coordinating work among multiple developers
-   Managing project versions
-   Supporting non-linear development (branching)

**Created by**: Linus Torvalds (2005)\
**Purpose**: Managing Linux kernel development\
**License**: GPL v2 (Free and Open Source)

------------------------------------------------------------------------

## 2. Core Git Concepts

### Three States of Files

1.  **Working Directory**: Where you edit files
2.  **Staging Area (Index)**: Files marked for next commit
3.  **Repository (.git directory)**: Committed and saved permanently

```         
Working Directory → (git add) → Staging Area → (git commit) → Repository
```

### Repository Types

-   **Local Repository**: On your machine
-   **Remote Repository**: Hosted server (GitHub, GitLab)
-   **Bare Repository**: No working directory (central repos)

------------------------------------------------------------------------

## 3. Git Environment Setup

### Installation

**Windows**: Download Git for Windows (includes Git Bash)\
**Mac**: `brew install git` or download from git-scm.com\
**Linux**: `sudo apt-get install git` (Ubuntu/Debian)

### Initial Configuration

``` bash
# Set username
git config --global user.name "Your Name"

# Set email
git config --global user.email "your.email@example.com"

# Check configuration
git config --list

# Set default editor
git config --global core.editor "vim"

# Set default branch name
git config --global init.defaultBranch main
```

------------------------------------------------------------------------

## 4. Basic Git Commands

### Initializing a Repository

``` bash
# Create new repository
git init

# Clone existing repository
git clone <repository-url>
git clone https://github.com/username/repo.git
```

### Checking Status

``` bash
# View status of files
git status

# Short status
git status -s
```

### Adding Files to Staging

``` bash
# Add specific file
git add filename.R

# Add all files
git add .

# Add all R files
git add *.R

# Add files interactively
git add -i
```

### Committing Changes

``` bash
# Commit with message
git commit -m "Your commit message"

# Commit all tracked files (skip staging)
git commit -a -m "Commit message"

# Amend last commit
git commit --amend -m "Updated message"
```

### Viewing History

``` bash
# View commit history
git log

# Compact view
git log --oneline

# View last 5 commits
git log -5

# View with graph
git log --graph --oneline --all

# View specific file history
git log -- filename.R
```

------------------------------------------------------------------------

## 5. Working with Remote Repositories

### Managing Remotes

``` bash
# View remote repositories
git remote -v

# Add remote repository
git remote add origin https://github.com/username/repo.git

# Remove remote
git remote remove origin

# Rename remote
git remote rename origin upstream
```

### Pushing Changes

``` bash
# Push to remote
git push origin main

# Push and set upstream
git push -u origin main

# Push all branches
git push --all

# Force push (use with caution!)
git push --force
```

### Pulling Changes

``` bash
# Fetch and merge changes
git pull origin main

# Fetch without merging
git fetch origin

# Pull with rebase
git pull --rebase origin main
```

------------------------------------------------------------------------

## 6. Branching and Merging

### Branch Management

``` bash
# Create new branch
git branch feature-branch

# Switch to branch
git checkout feature-branch

# Create and switch (shortcut)
git checkout -b feature-branch

# Modern way (Git 2.23+)
git switch feature-branch
git switch -c new-branch

# List branches
git branch
git branch -a  # include remote branches

# Delete branch
git branch -d feature-branch
git branch -D feature-branch  # force delete

# Delete remote branch
git push origin --delete feature-branch
```

### Merging

``` bash
# Merge branch into current branch
git merge feature-branch

# Merge without fast-forward
git merge --no-ff feature-branch

# Abort merge
git merge --abort
```

### Handling Merge Conflicts

``` bash
# 1. Git will mark conflicts in files
# 2. Open conflicted files and resolve manually
# 3. Add resolved files
git add resolved-file.R

# 4. Complete merge
git commit
```

### Rebasing

``` bash
# Rebase current branch onto main
git rebase main

# Interactive rebase (last 3 commits)
git rebase -i HEAD~3

# Continue after resolving conflicts
git rebase --continue

# Abort rebase
git rebase --abort
```

------------------------------------------------------------------------

## 7. Advanced Git Commands

### Viewing Differences

``` bash
# Changes in working directory
git diff

# Changes in staging area
git diff --staged

# Difference between branches
git diff main feature-branch

# Difference for specific file
git diff filename.R
```

### Undoing Changes

``` bash
# Discard changes in working directory
git restore filename.R
git checkout -- filename.R  # old way

# Unstage file
git restore --staged filename.R
git reset HEAD filename.R  # old way

# Revert a commit
git revert <commit-hash>

# Reset to previous commit
git reset --soft HEAD~1  # keep changes
git reset --mixed HEAD~1  # unstage changes
git reset --hard HEAD~1  # discard changes (⚠️ dangerous!)
```

### Stashing Changes

``` bash
# Save work temporarily
git stash

# List stashes
git stash list

# Apply latest stash
git stash apply

# Apply and remove stash
git stash pop

# Stash with message
git stash save "Work in progress"

# Clear all stashes
git stash clear
```

### Tagging

``` bash
# Create lightweight tag
git tag v1.0

# Create annotated tag
git tag -a v1.0 -m "Version 1.0 release"

# List tags
git tag

# Push tag to remote
git push origin v1.0

# Push all tags
git push --tags

# Delete tag
git tag -d v1.0
git push origin --delete v1.0
```

------------------------------------------------------------------------

## 8. Git Best Practices

### Commit Messages

**Good commit message format:**

```         
Short summary (50 chars or less)

Detailed explanation if needed. Wrap at 72 characters.
Explain what and why, not how.

- Bullet points are okay
- Use present tense: "Add feature" not "Added feature"
```

**Examples:**

``` bash
git commit -m "Add data validation function for user input"
git commit -m "Fix bug in date parsing for CSV imports"
git commit -m "Update README with installation instructions"
```

### Branching Strategy

**Common strategies:**

1.  **Git Flow**:
    -   `main`: Production code
    -   `develop`: Integration branch
    -   `feature/*`: New features
    -   `hotfix/*`: Emergency fixes
2.  **GitHub Flow**:
    -   `main`: Always deployable
    -   `feature/*`: All development
3.  **Trunk-Based Development**:
    -   `main`: Single integration branch
    -   Short-lived feature branches

------------------------------------------------------------------------

## 9. Git with R Programming

### Tracking R Projects

``` bash
# Initialize R project with Git
git init

# Create .gitignore for R
echo ".Rproj.user
.Rhistory
.RData
.Ruserdata
*.Rproj
.DS_Store" > .gitignore

# Add files
git add .
git commit -m "Initial commit of R project"
```

### Example .gitignore for R Projects

```         
# R specific
.Rproj.user
.Rhistory
.RData
.Ruserdata
*.Rproj

# Data files (if large)
*.csv
*.xlsx
data/

# Output files
*.pdf
*.png
figures/

# OS specific
.DS_Store
Thumbs.db
```

### Using Git with RStudio

RStudio has built-in Git integration:

1.  **Tools** → **Version Control** → **Project Setup**
2.  Use Git pane for staging, committing
3.  Visual diff viewer
4.  Branch management
5.  Push/pull buttons

------------------------------------------------------------------------

# Part 4: GitHub

## What is GitHub?

**GitHub** is a web-based platform for hosting Git repositories and
collaborating on code.

### Git vs GitHub

| Feature    | Git                  | GitHub                  |
|------------|----------------------|-------------------------|
| Type       | Version control tool | Hosting service         |
| Location   | Local computer       | Cloud-based             |
| Purpose    | Track changes        | Collaboration platform  |
| Access     | Command line/GUI     | Web interface + Git     |
| Created by | Linus Torvalds       | Chris Wanstrath, et al. |
| Owned by   | Linux community      | Microsoft               |

------------------------------------------------------------------------

## GitHub Features

### Core Features

1.  **Repository Hosting**: Store and manage Git repositories
2.  **Collaboration**: Pull requests, code reviews
3.  **Issue Tracking**: Bug reports, feature requests
4.  **Project Management**: Kanban boards, milestones
5.  **Actions**: CI/CD automation
6.  **Pages**: Host static websites
7.  **Wikis**: Documentation
8.  **Security**: Dependency scanning, secret detection

------------------------------------------------------------------------

## Working with GitHub

### Creating a Repository

**On GitHub:** 1. Click "New repository" 2. Enter repository name 3. Add
description (optional) 4. Choose public/private 5. Initialize with
README (optional) 6. Add .gitignore template 7. Choose license 8. Click
"Create repository"

**From Command Line:**

``` bash
# Create local repo
git init
git add .
git commit -m "Initial commit"

# Connect to GitHub
git remote add origin https://github.com/username/repo.git
git push -u origin main
```

### Cloning a Repository

``` bash
# Clone repository
git clone https://github.com/username/repo.git

# Clone specific branch
git clone -b branch-name https://github.com/username/repo.git

# Clone to specific folder
git clone https://github.com/username/repo.git my-folder
```

### Forking Workflow

**Fork** = Personal copy of someone else's repository

1.  Click "Fork" on GitHub
2.  Clone your fork

``` bash
git clone https://github.com/your-username/repo.git
cd repo
```

3.  Add upstream remote

``` bash
git remote add upstream https://github.com/original-owner/repo.git
```

4.  Keep fork updated

``` bash
git fetch upstream
git merge upstream/main
```

### Pull Requests

**Creating a Pull Request:**

1.  Create and push feature branch

``` bash
git checkout -b feature-branch
# make changes
git push origin feature-branch
```

2.  Go to GitHub repository
3.  Click "New pull request"
4.  Select base and compare branches
5.  Add title and description
6.  Click "Create pull request"

**Reviewing Pull Requests:** - Review code changes - Add comments -
Approve or request changes - Merge when approved

------------------------------------------------------------------------

## GitHub Best Practices

### README.md Template

``` markdown
# Project Title

Brief description of project

## Installation

\`\`\`r
install.packages("packagename")
\`\`\`

## Usage

\`\`\`r
library(packagename)
# example code
\`\`\`

## Contributing

Pull requests welcome!

## License

MIT License
```

### Contributing Guidelines

1.  **Fork** the repository
2.  **Create** feature branch (`git checkout -b feature/AmazingFeature`)
3.  **Commit** changes (`git commit -m 'Add some AmazingFeature'`)
4.  **Push** to branch (`git push origin feature/AmazingFeature`)
5.  **Open** Pull Request

------------------------------------------------------------------------

# Part 5: Git in CI/CD and Deployment

## Continuous Integration/Continuous Deployment

### What is CI/CD?

-   **CI**: Automatically test code when pushed
-   **CD**: Automatically deploy code when tests pass

### GitHub Actions Example

``` yaml
# .github/workflows/r-check.yml
name: R-CMD-check

on: [push, pull_request]

jobs:
  R-CMD-check:
    runs-on: ubuntu-latest
    steps:
      - uses: actions/checkout@v2
      - uses: r-lib/actions/setup-r@v2
      - name: Install dependencies
        run: |
          install.packages(c("remotes", "rcmdcheck"))
        shell: Rscript {0}
      - name: Check
        run: rcmdcheck::rcmdcheck(args = "--no-manual", error_on = "error")
        shell: Rscript {0}
```

------------------------------------------------------------------------

# Summary and Key Takeaways

## Control Flow

-   **if/else**: Decision making based on conditions
-   **for loop**: Known number of iterations
-   **while loop**: Condition-based repetition
-   **repeat loop**: Infinite loop with manual break
-   **switch**: Multiple case handling

## Version Control

-   **Local VCS**: Single machine, no collaboration
-   **Centralized VCS**: Single server, central authority (SVN)
-   **Distributed VCS**: Multiple copies, flexible workflow (Git)

## Git Essentials

``` bash
# Basic workflow
git init                    # Initialize
git add .                   # Stage changes
git commit -m "message"     # Commit
git push origin main        # Push to remote
git pull origin main        # Pull from remote

# Branching
git checkout -b feature     # Create branch
git merge feature           # Merge branch
git branch -d feature       # Delete branch
```

## GitHub Workflow

1.  **Fork** repository
2.  **Clone** to local
3.  **Branch** for features
4.  **Commit** changes
5.  **Push** to GitHub
6.  **Pull Request** for review
7.  **Merge** after approval

------------------------------------------------------------------------

# Practical Exercises

## Exercise 1: Control Flow

Create a grade calculator that: - Takes a score input - Uses if-else-if
to assign letter grade - Uses switch to assign GPA - Prints formatted
result

## Exercise 2: Loops

Write R code to: - Calculate Fibonacci sequence (for loop) - Find prime
numbers (while loop) - Create menu system (repeat loop)

## Exercise 3: Git Practice

1.  Create new R project
2.  Initialize Git repository
3.  Create multiple commits
4.  Create and merge feature branch
5.  Push to GitHub

## Exercise 4: Collaboration

1.  Fork a public R repository
2.  Clone to local
3.  Create feature branch
4.  Make improvements
5.  Submit pull request

------------------------------------------------------------------------

# Additional Resources

## Documentation

-   [Git Official Documentation](https://git-scm.com/doc)
-   [GitHub Guides](https://guides.github.com/)
-   [R for Data Science](https://r4ds.had.co.nz/)
-   [Happy Git with R](https://happygitwithr.com/)

## Cheat Sheets

-   [Git Cheat
    Sheet](https://education.github.com/git-cheat-sheet-education.pdf)
-   [GitHub Flow](https://guides.github.com/introduction/flow/)
-   [R Programming Cheat
    Sheet](https://cran.r-project.org/doc/contrib/Short-refcard.pdf)

## Online Tutorials

-   [Learn Git Branching](https://learngitbranching.js.org/)
-   [GitHub Learning Lab](https://lab.github.com/)
-   [Codecademy Git Course](https://www.codecademy.com/learn/learn-git)

------------------------------------------------------------------------

**End of Notes**

*Created with R Markdown* 📝
